Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-04-24 PAPERugh.tex
Tue Apr 24 18:10:09 EDT 2018

\newcommand{\cc}[1]{{\color{purple} \textbf{[CC: #1]}}}
\newcommand{\acl}[1]{{\color{red} \textbf{[ACL:  #1]}}}
\newcommand{\dcj}[1]{{\color{orange} \textbf{[DCJ: #1]}}}
\newcommand{\mjk}[1]{ {\color{olive} \textbf{[MJK: #1]}} }
\newcommand{\jea}[1]{{\color{blue} \textbf{[JEA: #1]}} }
\newcommand{\jp}[1]{ {\color{pink} \textbf{[JCP: #1]}} } %y'all picked all the good colors

@@

@@

@@

One important feature of weighting data by its true inverse covariance is that, while it can suppress some foregrounds, it cannot suppress the EoR signal (nor foreground modes that masquerade as signal-like modes). By construction, inverse covariance weighting in the quadratic estimator does not lead to signal loss. Therefore, in an ideal world with perfect foreground, 
instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data and use inverse covariance weighting to produce minimum variance power spectrum estimates without destroying the signal. In other words, the optimal quadratic estimator is by construction an unbiased estimator of the power spectrum.

@@

Suppose we have four independent measurements made along two different axes. As an example, we construct $\textbf{x}_{1a}$, $\textbf{x}_{1b}$, $\textbf{x}_{2a}$, and $\textbf{x}_{2b}$, where the numbers symbolize two different days of data and the letters represent two different baselines. Each of the 
measurements have dimensions of $100$ time integrations and $20$ frequency channels. They each have separate thermal 
noises constructed as a Gaussian random signal for each, and identical EoR signals.

To mimic the presence of a systematic, we add a toy sinusoid foreground, similar to the one used in 
Section \ref{sec:toymodel}, to only $\textbf{x}_{2a}$ and $\textbf{x}_{2b}$. This represents a foreground signal 
present in only the second day of data, but not the first. Mathematically, if  $\textbf{n}$ 
is noise, $\textbf{e}$ is the EoR signal, and $\textbf{f}$ is the foreground signal, the four measurements can be written as:

\begin{eqnarray}
\label{eq:bias1}
\textbf{x}_{1a} &=& \textbf{n}_{1a} + \textbf{e}  \\ 
\textbf{x}_{1b} &=& \textbf{n}_{1b} + \textbf{e}  \\ 
\textbf{x}_{2a} &=& \textbf{n}_{2a} + \textbf{e} + \textbf{f} \\ 
\textbf{x}_{2b} &=& \textbf{n}_{2b} + \textbf{e} + \textbf{f}.
\end{eqnarray}

\noindent We now take a jackknife along the day-axis, forming separate power spectrum estimates for each day:

\begin{eqnarray}
\widehat{\textbf{P}}_{1} &\propto& \textbf{x}_{1a}\textbf{x}_{1b}^{\dagger} \\
\widehat{\textbf{P}}_{2} &\propto& \textbf{x}_{2a}\textbf{x}_{2b}^{\dagger}.
\end{eqnarray}

@@

To construct a null test, we difference the two power spectra, with the result shown in Figure \ref{fig:toy_bias1} (black) along with the power spectrum of noise only (green). Subtracting the two estimates removes sky signal that should ideally be present in both jackknives. However, we see a clear difference between the null test and the power spectrum of 
noise. This signifies a non-EoR bias that is only present in either $\textbf{x}_{1}$ or $\textbf{x}_{2}$, but not both.

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias1.pdf}
	\caption{A null jackknife test shown as the power spectrum difference between two measurements (black), compared to the power spectrum of noise alone (green). Because the null test is not consistent with noise, it suggests the 
presence of a systematic in either $\textbf{x}_{1}$ or $\textbf{x}_{2}$. Null tests of clean measurements should be consistent 
with thermal noise.}
	\label{fig:toy_bias1}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias2.pdf}
	\caption{Power spectrum estimates for $\textbf{x}_{1}$ and $\textbf{x}_{2}$, two jackknives of the toy model. They suggest 
the presence of a systematic in $\textbf{x}_{2}$ only (which is exactly what was put in), illustrating how jackknives can be used to tease out excesses. Clean 
measurements should remain consistent despite the jackknife taken.}
	\label{fig:toy_bias2}
\end{figure}

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing 
which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $
\textbf{x}_{2}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and 
provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a 
likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving 
data before and after the application of a fringe-rate filter can reveal whether crosstalk noise bias is successfully suppressed 
with the filter, or if similar-shaped detections in both power spectra suggest otherwise. There are many other jackknife axes of 
which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist 
through them all and a clean measurement should exhibit noise-like null spectra.

@@

As a brief review, PAPER is a dedicated 21\,cm experiment located in the Karoo Desert in South Africa. The PAPER-64 
configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we 
focus solely on Stokes I estimated data \citep{moore_et_al2013} from PAPER's $30$ m East/West baselines (Figure 
\ref{fig:ant_layout}). All data is compressed, calibrated (using self-calibration and redundant calibration), delay-filtered (to remove foregrounds inside the wedge), LST-binned, and fringe-rate filtered. For detailed information about the backend system of PAPER-64, its observations, and data reduction pipeline, we 
refer the reader to \citet{parsons_et_al2010} and \citetalias{ali_et_al2015}. We note that all data processing steps are identical to those in \citetalias{ali_et_al2015} until after the LST-binning step in Figure 3 of \citetalias{ali_et_al2015}.

@@

We are interested in quantifying how much variance in $\textbf{e}$ is lost after weighting $\textbf{r}$ and estimating the power 
spectrum according to QE formalism. We investigate this by comparing two quantities we call the input power spectrum and 
output power spectrum: $\widehat{P}_{\rm in}$ and $\widehat{P}_{\rm out}$, estimated using QE as

\begin{equation}
\label{eq:Pin}
\widehat{P}_{\rm in,\alpha} \equiv \text{M}^{\alpha}_{\rm in}\textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
\widehat{P}_{\rm out,\alpha} &\equiv& \widehat{\textbf{P}}_{r,\alpha} \nonumber\\%-\widehat{\textbf{P}}_{x,\alpha} \nonumber \\
&=& \text{M}^{\alpha}_{r}\textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r},% - \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x},
\end{eqnarray}
where, for illustrative purposes and notational simplicity, we have written these equations with scalar normalizations M, even though for our numerical results we choose a non-diagonal matrix normalization using $\mathbf{M}$ as in Equation \eqref{eq:phat}.

The quantity $\widehat{P}_{\rm in}$, defined by Equation \eqref{eq:Pin}, is a uniformly weighted estimator of the power spectrum of $\mathbf{e}$. Since there is no noise contribution to $\mathbf{e}$, it can be considered the power spectrum of this particular realization of the EoR; alternatively, it can be viewed as the true power spectrum of the injected signal up to cosmic variance fluctuations. The role of $\widehat{P}_{\rm in}$ in our analysis is to serve as a reference for the power spectrum that would be measured if there were no signal loss or other systematics. This is then to be compared to $\widehat{P}_{\rm out}$, which approximates the (lossy) power spectrum estimate that is output by our analysis pipeline prior to any signal loss adjustments. In principle, one could compute the true $\widehat{P}_{\rm out}$ using end-to-end simulations of the instrument and data analysis pipeline. In practice, however, such simulations may not accurately reflect real-life systematics and foregrounds. To overcome this obstacle, one can make the assumption that since the EoR signal is expected to be small, the data vector $\mathbf{x}$ itself is our best model of these contaminants. Making this assumption, the injected EoR signal $\mathbf{e}$ takes on the role of the true EoR signal, and the sum of $\mathbf{x}$ and $\mathbf{e}$ (i.e., Equation \eqref{eq:rxe}) replaces $\mathbf{x}$ as our model of the measured data. Therefore, $\widehat{P}_{\rm out}$ can be directly compared to the measurement that we make.

@@

%\begin{equation}
%\widehat{P}_{\rm out} = \widehat{\textbf{P}}_{x} \label{eqn:anzatz},
%\end{equation}
%and folding in our definition from Equation \eqref{eq:sigloss}, this ansatz implies that we are seeking the injected 
%signal amplitude which causes a power doubling ($\phat_r = 2 \phat_x$). \cc{I am unsure of this power doubling thing}

Under this injection framework, we can begin to see explicitly why there can be large signal loss. Expanding out Equation \eqref{eq:sigloss}, $\widehat{P}_{\rm out}$ becomes:

@@

%In short, we can approximate a signal loss estimate as the ratio of $\widehat{P}_{\rm out}/P_{\rm in}$, evaluated at the data level $\widehat{\textbf{P}}_{x}$. We motivate the fact that we can evaluate the output-to-input power spectrum ratio at $\widehat{\textbf{P}}_{x}$ by the following reasoning (and then detail our signal loss estimation in practice in the sections that follow).

@@

%Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $\widehat{P}_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our QE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross-terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross-correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}

@@

%Effectively, this means that the relationship between the input and output power spectra, $P_{\rm in}$ and $\widehat{P}_{\rm out}$, can be thought of as a transfer function 
%which, for a sampling of $P_{\rm in}$ and $\widehat{P}_{\rm out}$ provides a mapping from an input power spectrum distribution into an output 

@@

%\subsubsection{Signal Loss Illustration}
%\label{sec:Illustration}
%To explore how our expression for $\widehat{P}_{\rm out}$ encapsulates signal loss, we expand out Equation \eqref{eq:sigloss}:

\begin{eqnarray}
\label{eq:crossterm_full}
\widehat{P}_{\rm out,\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) \nonumber \\%- 
%\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms (terms with one copy of $\textbf{e}$ and one copy of $\textbf{x}$) can be summed together as:

\begin{eqnarray}
\label{eq:crossterm}
\widehat{P}_{\rm out,\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: empirically estimated inverse covariance weighting ($\textbf{R}_{r} \equiv \widehat{\textbf{C}}_{r}^{-1}$) and uniform weighting ($\textbf{R}_{r} \equiv \textbf{I}$). We will now examine the behavior of this equation in three different regimes of the injected signal - very weak (left ends of the $P_{\rm in}$ axes in Figure \ref{fig:sigloss_terms}), very strong (right ends), and in between (middle portions).

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are Gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniformly weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a symmetric distribution with zero mean and width $\sqrt{\widehat{\textbf{P}}_e}\sqrt{\widehat{\textbf{P}}_x}$. We also note that in this regime, $\widehat{\textbf{P}}_{r}$ (black) approaches the data-only power spectrum value (gray) as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $\widehat{P}_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $\widehat{P}_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

>>>>

\newcommand{\cc}[1]{{\color{purple} \textbf{[CC: #1]}}}
\newcommand{\acl}[1]{{\color{red} \textbf{[ACL:  #1]}}}
\newcommand{\dcj}[1]{{\color{orange} \textbf{[DCJ: #1]}}}
\newcommand{\mjk}[1]{ {\color{olive} \textbf{[MJK: #1]}} }
\newcommand{\jea}[1]{{\color{blue} \textbf{ #1}} }
\newcommand{\jp}[1]{ {\color{pink} \textbf{[JCP: #1]}} } %y'all picked all the good colors

@@

% First, explain the estimator and what it is supposed to do.
% Then show that using empirical covariance naively will lead to problems.
% Then discuss what happens with not enough modes - here the convergence to the correct eigenvectors is discussed

@@

\jea{One important feature of weighting data by its true inverse covariance is that, while it can suppress some foregrounds, it cannot suppress the EoR signal (nor foreground modes that masquerade as signal-like modes). By construction, inverse covariance weighting in the quadratic estimator does not lead to signal loss. Therefore, in an ideal world with perfect foreground, 
instrumental, and EoR models, we could form $\textbf{C}$ in a way that accurately describes our measured data and use inverse covariance weighting to produce minimum variance power spectrum estimates without destroying the signal. In other words, the optimal quadratic estimator is by construction an unbiased estimator of the power spectrum.}

@@

@@

Suppose we have two measurements (for example, from two baselines), $\textbf{x}_{a}$ and $\textbf{x}_{b}$. The 
measurements have dimensions of $200$ time integrations and $20$ frequency channels. They each have separate thermal 
noises constructed as a Gaussian random signal for each, and identical EoR signals.

To mimic the presence of a systematic in part of the measurement, we add a toy sinusoid foreground, similar to the one used in 
Section \ref{sec:toymodel}, to the first $100$ time integrations of both measurements. This represents a foreground signal 
present in, for example, the first half of the LST range used for analysis, but not the second half. Mathematically, if  $\textbf{n}$ 
is noise, $\textbf{e}$ is the EoR signal, and $\textbf{f}$ is the foreground signal, the two measurements (which are cross-multiplied to form power spectra) can be written as:

\begin{equation}
\label{eq:bias1}
\textbf{x}_{a} = \textbf{n}_{a} + \textbf{e} + \textbf{f}
\end{equation}

\begin{equation}
\label{eq:bias2}
\textbf{x}_{b} = \textbf{n}_{b} + \textbf{e} + \textbf{f}.
\end{equation}

\noindent The two jackknife samples are $\textbf{x}_{1}$ and $\textbf{x}_{2}$, representing jackknives in time. These can be 
written (for both measurements $a$ and $b$) as:

\begin{eqnarray}
\textbf{x}_{1} &=& \textbf{n}_{1} + \textbf{e}_{1} + \textbf{f} \\
\textbf{x}_{2} &=& \textbf{n}_{2} + \textbf{e}_{2}
\end{eqnarray}

\noindent For example, $\textbf{x}_{1a}$ represents a jackknife sample (first half of the data) for the first measurement. 
Similarly, $\textbf{x}_{2b}$ represents a jackknife sample (second half of the data) for the second measurement. We note that $\textbf{e}$ itself will be different in two jackknives taken along LST, but could be identical in other cases (such as a jackknife along even/odd Julian days).

@@

We form three different power spectrum estimates shown in Figure \ref{fig:toy_bias1}. The first is a null test where we subtract $
\textbf{x}_{2}$ from $\textbf{x}_{1}$ for both measurements $a$ and $b$. This is equivalent to splitting up a full dataset along 
an axis (in this case, time) and subtracting the two to remove sky signal that should ideally be present in both. We cross-multiply measurements $a$ and $b$ to form an un-biased (thermal noise-wise) estimate (blue curve). The second estimate, 
shown in red, is the same null test with the foreground systematic removed (eliminate $\textbf{f}$ in Equations \ref{eq:bias1} 
and \ref{eq:bias2}). Finally, we also show the noise power spectrum (green).

From this test we see a clear difference between the null test with the presence of the foregrounds, and the power spectrum of 
noise. This signifies a non-EoR bias that is only present in either $\textbf{x}_{1}$ or $\textbf{x}_{2}$, but not both.

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias1.pdf}
	\caption{Power spectrum estimates for a null jackknife test with the presence of a foreground systematic (blue), without 
the foreground systematic (red), and noise alone (green). Because the first null test is not consistent with noise, it suggests the 
presence of a systematic in either $\textbf{x}_{1}$ or $\textbf{x}_{2}$. Null tests of clean measurements should be consistent 
with thermal noise.}
	\label{fig:toy_bias1}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias2.pdf}
	\caption{Power spectrum estimates for $\textbf{x}_{1}$ and $\textbf{x}_{2}$, two jackknives of the toy model. They suggest 
the presence of a systematic in $\textbf{x}_{1}$ only (which is exactly what was put in), illustrating how jackknives can be used to tease out excesses. Clean 
measurements should remain consistent despite the jackknife taken.}
	\label{fig:toy_bias2}
\end{figure}

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing 
which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $
\textbf{x}_{1}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and 
provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a 
likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving 
data before and after the application of a fringe-rate filter can reveal whether crosstalk noise bias is successfully suppressed 
with the filter, or if similar-shaped detections in both power spectra suggest otherwise. There are many other jackknife axes of 
which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist 
through them all and a clean measurement should exhibit noise-like null spectra.

@@

As a brief review, PAPER is a dedicated 21\,cm experiment located in the Karoo Desert in South Africa. The PAPER-64 
configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we 
focus solely on Stokes I estimated data (rather than the full Stokes considered in 
\citep{moore_et_al2017})
%JEA - I didn't understand the original citation to Moore 2013
from PAPER's $30$ m East/West baselines (Figure 
\ref{fig:ant_layout}). All data is compressed, calibrated (using self-calibration and redundant calibration), delay-filtered (to remove foregrounds inside the wedge), LST-binned, and fringe-rate filtered. For detailed information about the backend system of PAPER-64, its observations, and data reduction pipeline, we 
refer the reader to \citet{parsons_et_al2010} and \citetalias{ali_et_al2015}. We note that all data processing steps are identical to those in \citetalias{ali_et_al2015} until after the LST-binning step in Figure 3 of \citetalias{ali_et_al2015}.

@@

We are interested in quantifying how much variance in $\textbf{e}$ is lost after weighting $\textbf{r}$ and estimating the power 
spectrum according to QE formalism. We investigate this by comparing two quantities we call the input power spectrum and 
output power spectrum: $P_{\rm in}$ and $P_{\rm out}$, defined as

\begin{equation}
\label{eq:Pin}
P_{\rm in,\alpha} \equiv \text{M}^{\alpha}_{\rm in}\textbf{e}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{\rm out,\alpha} &\equiv& \widehat{\textbf{P}}_{r,\alpha} \nonumber\\%-\widehat{\textbf{P}}_{x,\alpha} \nonumber \\
&=& \text{M}^{\alpha}_{r}\textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r},% - \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x},
\end{eqnarray}
where, for illustrative purposes and notational simplicity, we have written these equations with scalar normalizations M, even though for our numerical results we choose a non-diagonal matrix normalization using $\mathbf{M}$ as in Equation \eqref{eq:phat}.

The quantity $P_{\rm in}$, defined by Equation \eqref{eq:Pin}, is a uniformly weighted estimator of the power spectrum of $\mathbf{e}$. Since there is no noise contribution to $\mathbf{e}$, it can be considered the power spectrum of this particular realization of the EoR; alternatively, it can be viewed as the true power spectrum of the injected signal up to cosmic variance fluctuations. The role of $P_{\rm in}$ in our analysis is to serve as a reference for the power spectrum that would be measured if there were no signal loss or other systematics. This is then to be compared to $P_{\rm out}$, which approximates the (lossy) power spectrum estimate that is output by our analysis pipeline prior to any signal loss adjustments. In principle, one could compute $P_{\rm out}$ using end-to-end simulations of the instrument and data analysis pipeline. In practice, however, such simulations may not accurately reflect real-life systematics and foregrounds. To overcome this obstacle, one can make the assumption that since the EoR signal is expected to be small, the data vector $\mathbf{x}$ itself is our best model of these contaminants. Making this assumption, the injected EoR signal $\mathbf{e}$ takes on the role of the true EoR signal, and the sum of $\mathbf{x}$ and $\mathbf{e}$ (i.e., Equation \eqref{eq:rxe}) replaces $\mathbf{x}$ as our model of the measured data. Therefore, $P_{\rm out}$ can be directly compared to the measurement that we make.

@@

%\begin{equation}
%P_{\rm out} = \widehat{\textbf{P}}_{x} \label{eqn:anzatz},
%\end{equation}
%and folding in our definition from Equation \eqref{eq:sigloss}, this ansatz implies that we are seeking the injected 
%signal amplitude which causes a power doubling ($\phat_r = 2 \phat_x$). \cc{I am unsure of this power doubling thing}

Under this injection framework, we can begin to see explicitly why there can be large signal loss. Expanding out Equation \eqref{eq:sigloss}, $P_{\rm out}$ becomes:

@@

%In short, we can approximate a signal loss estimate as the ratio of $P_{\rm out}/P_{\rm in}$, evaluated at the data level $\widehat{\textbf{P}}_{x}$. We motivate the fact that we can evaluate the output-to-input power spectrum ratio at $\widehat{\textbf{P}}_{x}$ by the following reasoning (and then detail our signal loss estimation in practice in the sections that follow).

@@

%Using this motivation, the numerator of Equation \eqref{eq:Deriv2} is precisely our expression for $P_{\rm out}$ (Equation \eqref{eq:sigloss}), and the denominator is $P_{\rm in}$ (Equation \eqref{eq:Pin}), where the function $g$ is our QE power spectrum pipeline. \dcj{Here is the post-hoc justification for this signal loss method.  Except goes back to assumption of equation for g which just \emph{assumes} that there are no cross-terms in the estimation of a power spectrum. So we're just kicking the can down the road.} \acl{Ok, in my defense, Equation \eqref{eq:LinearPspecSum} is a little better than an \emph{assumption}. Appendix \eqref{sec:sigloss_appendix} does in fact illustrate this with an example. The crucial question is whether we're allowed to motivate our method based on a relationship that is only true in expectation. Note that this expectation includes allowances for signal loss due to the cross-correlation between signal and foregrounds, with an example of this being Appendix \ref{sec:sigloss_appendix}.}

@@

%Effectively, this means that the relationship between the input and output power spectra, $P_{\rm in}$ and $P_{\rm out}$, can be thought of as a transfer function 
%which, for a sampling of $P_{\rm in}$ and $P_{\rm out}$ provides a mapping from an input power spectrum distribution into an output 

@@

%\subsubsection{Signal Loss Illustration}
%\label{sec:Illustration}
%To explore how our expression for $P_{\rm out}$ encapsulates signal loss, we expand out Equation \eqref{eq:sigloss}:

\begin{eqnarray}
\label{eq:crossterm_full}
P_{\rm out,\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) \nonumber \\%- 
%\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms (terms with one copy of $\textbf{e}$ and one copy of $\textbf{x}$) can be summed together as:

\begin{eqnarray}
\label{eq:crossterm}
P_{\rm out,\alpha} &= &  \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& 2 \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
In order to investigate the effect of each of these terms on signal loss, all three components are plotted in Figure \ref{fig:sigloss_terms} for two cases: empirically estimated inverse covariance weighting ($\textbf{R}_{r} \equiv \widehat{\textbf{C}}_{r}^{-1}$) and uniform weighting ($\textbf{R}_{r} \equiv \textbf{I}$). We will now examine the behavior of this equation in three different regimes of the injected signal - very weak (left ends of the $P_{\rm in}$ axes in Figure \ref{fig:sigloss_terms}), very strong (right ends), and in between (middle portions).

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are Gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniformly weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a symmetric distribution with zero mean and width $\sqrt{P_e}\sqrt{P_x}$. We also note that in this regime, $\widehat{\textbf{P}}_{r}$ (black) approaches the data-only power spectrum value (gray) as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $P_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $P_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the estimated inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a symmetric distribution of zero mean and width $\sqrt{P_e}\sqrt{P_x}$.

{\bf In between:}
When the injected signal is of a similar amplitude to the data by itself, the situation becomes less straightforward. We see that 
the weighted injected power spectrum component mirrors the input power indicating little loss (i.e. the green curve follows the dotted black line), eventually 
departing from unity when the injected amplitude is well above the level of the data power spectrum. However, 
in this regime the cross-term has nearly the same amplitude, but with a negative sign. As explained below, this negativity is the result of cross-correlating inverse covariance weighted terms.  This negative component drives down the $P_{\rm out}$ estimator (black). We note that in \citetalias{ali_et_al2015}, signal loss was computed by only looking at the second term in Equation \eqref{eq:crossterm} (green), which incorrectly implies no loss at the data-only power spectrum level. Ignoring the effect of the negative power from the cross-terms is the main reason for under-estimating power spectrum limits in \citetalias{ali_et_al2015}.
